The value of being indispensable: a cooperative approach to bargaining with a continuum of players

### 2023-11-28 src/paper/theory.tex
Add proposition for weighted values in the example
Add idiosyncratic utilities

\paragraph{Example: power function.}

@@

\begin{lemma}
    % TODO: check notation!
    \label{lem:many_sided_manifold}
    Assume that $X_n \xrightarrow[]{d} X$ where $X$ is a degenerate distribution in the following sense: $\exists \, a_l: [0, 1] \to [0, 1], l = 1, \dots, L$, such that
    \begin{align*}
        X = (a_1(\xi), \dots, a_L(\xi)),
    \end{align*}
    where $\xi$ is a random variable on $[0, 1]$ with cdf $H(s)$.
    In words, the whole probability mass is concentrated on the manifold $(a_1(s), \dots a_L(s)), t \in [0, 1]$.
    Then
    \begin{align*}
        \varphi_P^\infty = \lim_{n \to \infty} \varphi_P^n &= \int_0^1 f(a_1(s), \dots a_L(s)) \mathrm{d}H(s).
    \end{align*}
    Furthermore, if $f$'s partial derivatives exist on $\{(a_1(t), \dots, a_L(t)) : t \in [0, 1]\}$ the value of fringe $l$ has the following limit:
    \begin{align*}
        \varphi_{A^l}^\infty = \lim_{n \to \infty} \varphi_{A^l}^n &= \int_0^1 H(s) \partial_l f(a_1(s), \dots a_L(s)) \ds.        
    \end{align*}
\end{lemma}

@@

\begin{proposition}
    \label{prop:many_sided_weighted}
    Let $f$ be continuous on $[0, 1]^L$.
    Then,
    \begin{align*}
        \varphi_P^\infty & = \int_0^1 \lambda_P t^{\lambda_P - 1} f(t^{\lambda_1}, \dots, t^{\lambda^L}) \dt
    \end{align*}
    and
    \begin{align*}
        \varphi_{A^l}^\infty & = \int_0^1 t^{\lambda_P} \partial_l f(t^{\lambda_1}, \dots, t^{\lambda^L}) \dt.
    \end{align*}
\end{proposition}

@@

Following \textcite{armstrong2006competition}, let us also model player entry in a reduced form manner.
Assume that there is a continuum of potential entrants on both sides, and the number of actual entrants is an increasing function of the utility achieved after entering\footnote{
    One possible microfoundation of this is assuming some idiosyncratic entry cost for the players.
    Then, the number of entrants is the number of players for whom the utility of entering is higher than the entry cost.
}:
\begin{align*}
    n_i = \phi_i(u_i).
\end{align*}

@@

As an immediate corollary, utilities in this case are given by
\begin{align*}
    u_i^* = \alpha_i n_j ^ {\gamma_i} + \alpha_j \gamma_j n_j n_i^{\gamma_j - 1} - F_i.
\end{align*}
Furthermore, the utility of type $i$ depends positively on both $\alpha_i$ (mechanically) and $\alpha_j$ (a stronger positive externality on the other players implies a lower entry fee and thus higher utility).
Finally, if $n_i \geq e^{-\frac{1}{\gamma_j}}$, then $u_i$ is also increasing in $\gamma_j$.
Both of the latter two effects are related to player $i$ being compensated for the positive externality it imposes on the other side.

@@

For ease of comparison with the bargaining case, let us also derive the utilities in this case:
\begin{align*}
    u_i^u = \alpha_i n_j ^ {\gamma_i} + \alpha_j \gamma_j n_j n_i^{\gamma_j - 1} - F_i - \frac{\phi_1(u_1)}{\phi'_1(u_1)}.
\end{align*}
Clearly, as equilibrium utilities must be smaller than in the welfare-maximizing case, due to $\phi_i$ being increasing, the number of entrants is also lower.

@@

Finally, let us consider the case when the platform and the players bargain over the entry fees.
To do this, let us impose some structure on the timing of the game.
The idea is that even though the platform cannot commit to a specific entry fee structure anymore, it can commit to a bargaining process.
Specifically, it is able to commit to a bargaining weight for each side and itself: $\lambda_P, \lambda_1, \lambda_2$.

@@

The characteristic function of the game takes a similar form to that in \cref{sec:many_sided}.
If the platform is not part of the coalition, then the value is zero.
Otherwise, it is simply the sum of the utilities of the two sides minus the cost of serving them.\footnote{
    The entry fees do not matter, as they are simply payments between coalition members, and thus do not affect the total value.
}
This value, as a function of the number of entrants, is
\begin{align*}
    f(n_1, n_2) = n_1 \alpha_1 n_2^{\gamma_1} + n_2 \alpha_2 n_1^{\gamma_2} - F_1 n_1 - F_2 n_2.
\end{align*}

The share received by each type of player is then easily established using \cref{prop:many_sided_weighted}.
\begin{proposition}
    \label{prop:platform_bargaining_last_period}
    \begin{align*}
        \pi_P(n_1, n_2) &= \int_0^1 \lambda_P t^{\lambda_P - 1} f(t_1^{\lambda_1} n_1, t_1^{\lambda_2} n_2) \dt \\
        u_i(n_1, n_2) &= \int_0^1 t^{\lambda_P} \partial_i f(t_1^{\lambda_1} n_1, t_1^{\lambda_2} n_2) \dt
    \end{align*}
\end{proposition}

>>>>

\paragraph{Example: power function.}
\label{sec:power_function_example}

@@

\begin{lemma}
    % TODO: check notation!
    \label{lem:many_sided_manifold}
    Assume that $X_n \xrightarrow[]{d} X$ where $X$ is a degenerate distribution in the following sense: $\exists \, a_l: [0, 1] \to [0, 1], l = 1, \dots, L$, such that
    \begin{align*}
        X = (a_1(\xi), \dots, a_L(\xi)),
    \end{align*}
    where $\xi$ is a random variable on $[0, 1]$ with cdf $H(s)$.
    In words, the whole probability mass is concentrated on the manifold $(a_1(s), \dots a_L(s)), t \in [0, 1]$.
    Then
    \begin{align*}
        \varphi_P^\infty = \lim_{n \to \infty} \varphi_P^n &= \int_0^1 f(a_1(s), \dots a_L(s)) \mathrm{d}H(s).
    \end{align*}
    Furthermore, if $f$'s partial derivatives exist on $\{(a_1(t), \dots, a_L(t)) : t \in [0, 1]\}$ the value of fringe $l$ has the following limit:
    \begin{align*}
        \varphi_{A^l}^\infty = \lim_{n \to \infty} \varphi_{A^l}^n &= \int_0^1 H(s) a_l'(s) \partial_l f(a_1(s), \dots a_L(s)) \ds.        
    \end{align*}
\end{lemma}

@@

\begin{proposition}
    \label{prop:many_sided_weighted}
    Let $f$ be continuous on $[0, 1]^L$.
    Then,
    \begin{align*}
        \varphi_P^\infty & = \int_0^1 \lambda_P t^{\lambda_P - 1} f(t^{\lambda_1}, \dots, t^{\lambda^L}) \dt
    \end{align*}
    and
    \begin{align*}
        \varphi_{A^l}^\infty & = \int_0^1 t^{\lambda_P} \lambda_l t^{\lambda_l - 1} \partial_l f(t^{\lambda_1}, \dots, t^{\lambda^L}) \dt.
    \end{align*}
\end{proposition}

@@

Following \textcite{armstrong2006competition}, let us also model player entry in a reduced form manner.
There is a continuum of potential entrants on both sides of the market.
Assume that the total utility of a player on side $i$ also includes an idiosyncratic component.
Let us order the players on side $i$ by this idiosyncratic utility in a decreasing manner.
That is, if we denote the idiosyncratic utility of player $n$ on side $i$ as $\epsilon_i(n)$, then $\epsilon_i$ is strictly decreasing in $n$.
Also assume that $\lim_{n \to \infty} \epsilon_i(n) = -\infty$.

Entry is given by the usual assumption that a player enters if and only if the total utility of entering is non-negative.

Due to the properties of $\epsilon_i$, this implies that there is a well-defined, finite number of entrants on each side.
Let us denote the number of entrants as a function of their non-idiosyncratic utility as $\phi_i(u_i)$.
Then, the total number of players on side $i$ is 
\begin{align*}
    n_i = \phi_i(u_i) = \epsilon_i^{-1}(u_i),
\end{align*}
where $\phi_i$ has the same role as in \textcite{armstrong2006competition}.

@@

As an immediate corollary, utilities in this case are given by
\begin{align*}
    u_i^*(n_i, n_j) = \alpha_i n_j ^ {\gamma_i} + \alpha_j \gamma_j n_j n_i^{\gamma_j - 1} - F_i
\end{align*}
for an individual player on side $i$.
That is, side $i$ obtains the total utility generated by its players minus the cost of serving them, plus $\gamma_j$ times the total utility generated by the other side, too. This is due to the positive externality that side $i$ imposes on side $j$.

Furthermore, the utility of type $i$ depends positively on both $\alpha_i$ (mechanically) and $\alpha_j$ (a stronger positive externality on the other players implies a lower entry fee and thus higher utility).
Finally, if $n_i \geq e^{-\frac{1}{\gamma_j}}$, then $u_i$ is also increasing in $\gamma_j$.
Both of the latter two effects are related to player $i$ being compensated for the positive externality it imposes on the other side.

One can also obtain the total welfare of side $i$, as it is simply the number of entrants plus the non-idiosyncratic utility of its players, plus the sum (integral) of the idiosyncratic utilities.
\begin{align*}
    w_i^*(n_i, n_j) = \alpha_i n_i n_j ^ {\gamma_i} + \gamma_j \alpha_j n_j n_i^{\gamma_j} - n_i F_i + \int_0^{n_i} \epsilon_i(s) \ds.
\end{align*}

@@

For ease of comparison with the bargaining case, let us also derive the utilities in this case:
\begin{align*}
    u_i^u(n_1, n_2) &= \alpha_i n_j ^ {\gamma_i} + \alpha_j \gamma_j n_j n_i^{\gamma_j - 1} - F_i - \frac{\phi_1(u_i)}{\phi'_1(u_i)} \\
    &= \alpha_i n_j ^ {\gamma_i} + \alpha_j \gamma_j n_j n_i^{\gamma_j - 1} - F_i - n_i \epsilon'_i(n_i).
\end{align*}
Welfare of side $i$ immediately follows in this case, too:
\begin{align*}
    U_i^u(n_1, n_2) = \alpha_i n_i n_j ^ {\gamma_i} + \gamma_j \alpha_j n_j n_i^{\gamma_j} - n_i F_i - n_i^2 \epsilon'_i(n_i) + \int_0^{n_i} \epsilon_i(s) \ds.
\end{align*}
Clearly, equilibrium utilities must be smaller than in the welfare-maximizing case, due to $\phi_i$ being increasing, the number of entrants is also lower.

@@

Finally, let us consider the case when the platform and the players bargain over the value they generate.
To do this, let us impose some structure on the timing of the game.
The idea is that even though the platform cannot commit to a specific entry fee structure anymore, it can commit to a bargaining process.
Specifically, it is able to commit to a bargaining weight for each side and itself: $\lambda_P, \lambda_1, \lambda_2$.

@@

The characteristic function of the game takes a similar form to that in \cref{sec:many_sided}.
If the platform is not part of the coalition, then the value is zero.
Otherwise, it is simply the sum of the utilities of the two sides minus the cost of serving them.\footnote{
    The entry fees do not matter, as they are simply payments between coalition members, and thus do not affect the total value.
}
This value, as a function of the number of entrants, is
\begin{align*}
    w(n_1, n_2) = n_1 \alpha_1 n_2^{\gamma_1} + n_2 \alpha_2 n_1^{\gamma_2} - F_1 n_1 - F_2 n_2.
\end{align*}

The share received by each type of player is then easily established using \cref{prop:many_sided_weighted}.
\begin{proposition}
    \label{prop:platform_bargaining_last_period}
    Let $f(t_1, t_2) = w(t_1 n_1, t_2 n_2)$. Then the weighted values of the various sides are
    \begin{alignat*}{4}
        \pi_P^b(n_1, n_2) &= \underbracket{\frac{\lambda_P}{\lambda_P + \lambda_1 + \lambda_2\gamma_1}}_{\coloneqq S_P^{b_1}} \alpha_1 n_1 n_2^{\gamma_1} &&+ \underbracket{\frac{\lambda_P}{\lambda_P + \lambda_2 + \lambda_1\gamma_2}}_{\coloneqq S_P^{b_2}} \alpha_2 n_2 n_1^{\gamma_2} &&- \underbracket{\frac{\lambda_P}{\lambda_P + \lambda_1}}_{\coloneqq S_P^{c_1}} n_1 F_1 &&- \underbracket{\frac{\lambda_P}{\lambda_P + \lambda_2}}_{\coloneqq S_P^{c_2}} n_2 F_2 \\
        U_1^b(n_1, n_2) &= \underbracket{\frac{\lambda_1}{\lambda_P + \lambda_1 + \lambda_2\gamma_1}}_{\coloneqq S_1^{b_1}} \alpha_1 n_1 n_2^{\gamma_1} &&+ \underbracket{\frac{\lambda_1 \gamma_2}{\lambda_P + \lambda_2 + \lambda_1\gamma_2}}_{\coloneqq S_1^{b_2}}  \alpha_2 n_2 n_1^{\gamma_2} &&- \underbracket{\frac{\lambda_1}{\lambda_P + \lambda_1}}_{\coloneqq S_1^{c_1}} n_1 F_1 && \\
        U_2^b(n_1, n_2) &= \underbracket{\frac{\lambda_2\gamma_1}{\lambda_P + \lambda_1 + \lambda_2\gamma_1}}_{\coloneqq S_2^{b_1}} \alpha_1 n_1 n_2^{\gamma_1} &&+ \underbracket{\frac{\lambda_2}{\lambda_P + \lambda_2 + \lambda_1\gamma_2}}_{\coloneqq S_2^{b_1}} \alpha_2 n_2 n_1^{\gamma_2} && &&- \underbracket{\frac{\lambda_2}{\lambda_P + \lambda_2}}_{\coloneqq S_2^{c_2}} n_2 F_2.
    \end{alignat*}
\end{proposition}

There are four parts that the total welfare $w(n_1, n_2)$ is divided into: utility generated by side $1$ and $2$, and cost of serving side $1$ and $2$.
Each side's share of these parts is given by $S_i^{b_1}, S_i^{b_2}, S_i^{c_1}$ and $S_i^{c_2}$, respectively.
As it turns out, these shares only depend on the bargaining weights ($\lambda_P, \lambda_1, \lambda_2$) and the network effects ($\gamma_1, \gamma_2$), but not on the actual number of entrants.\footnote{
    This is due to the characteristic function being a polynomial in the number of entrants.
    Just like in \cref{sec:power_function_example}, the parts of the polynomial are divided in a constant proportion.
}
Furthermore, all of these shares are increasing in one's own bargaining weight, and decreasing in the others' weights.
Finally, the share of side $i$ from the value generated on side $j$ is also increasing in $\gamma_j$.
This means that if players on side $i$ are less substitutable / more complementary in terms of generating utility on side $j$, then they will receive a larger share of the value generated on that side.

Let us look at the per-player utilities to
